validateHOT - Validate your Holdout Task

### 2023-10-28 paper/paper.Rmd
paper update - 231028

validateHOT is a package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using [Sawtooth Software](https://sawtoothsoftware.com/).

Preference measurement techniques', such as (A)CBC or MaxDiff, ultimate goal is to predict future behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and can also predict outside tasks. The easiest way to test this is to include so-called holdout or validation task [@Orme2015]. Despite the important role of validation tasks, practictioners often do not include them [@yang2018], which is unsatisfactory, given the fact that the model is used to estimate market shares which is the basis for relevant marketing decisions.

validateHOT combines both validation and market simulation in one package. validateHOT's advantages are the following: a) it helps you to decide which is the best model to proceed by validating it, b) it runs relevant market simulations that help to find the right product combinations, and finally, c) it is open source tool for functions that are usually implemented in Sawtooth Software and are often a black-box for researchers and practitioners.

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and therefore ensuring that the model has predictive validity; b) running market simulations. Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix (see \autoref{tab:table1}). However, to put the Sawtooth export into the right format, the user needs some data wrangling which can be a barrier. Other packages mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR [@ChoiceModelR], choicetools [@choicetools], logitR [@logitr], bayesm [@bayesm] etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations is still missing. The goal of validateHOT is still to fill this research gap. validateHOT creates market shares based on *share of preferences* or *first choice rule*, which can be helpful to report results for research articles according to open science standards. In addition, it also provides the function \texttt{\color{purple}turf}, which was offered formerly by the turfR package, which is no longer available on the Comprehensive R Archive Network (see [CRAN](https://cran.r-project.org/web/packages/turfR/index.html)). So, currently practitioners and academics mainly have to stick to paid solutions, for example, Sawtooth Software.

@@

The main function that creates the validation/holdout task is the \texttt{\color{purple}createHOT} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. The other functions in the package mainly focus on four different components which are outlined in \autoref{tab:table1}.

| Validation metrics | Confusion matrix | Market Simulations | Rescaling scores |
|:----------------:|:----------------:|:----------------:|:----------------:|
|     hitrate()      |    accuracy()    |    freqassort()    |    att_imp()     |
|        kl()        |       f1()       |     marksim()      |  prob_scores()   |
|       mae()        |   precision()    |      reach()       |                  |
|      medae()       |     recall()     |       turf()       |                  |
|       mhp()        |  specificity()   |                    |                  |
|       rmse()       |                  |                    |                  |

@@

After running the Hierarchical Bayes estimation in Sawtooth Software, the **raw** utility scores have to be exported and afterwards read into *R* including the actual choice in the validation/holdout task.

To define the validation/holdout task, which has a total of 7 items (\texttt{\color{purple}prod}) plus the no-buy alternative (\texttt{\color{purple}none}), we use the \texttt{\color{purple}createHOT} function. Here, the user can define the attributes as well as the method (in this case \texttt{\color{purple}MaxDiff}).

@@

Next, to get the relevant validation metrics that are often reported in conjoint studies, for example, hit rate [e.g., @ding2005], mean hit probability [mhp, @voleti2017], or mean absolute error [mae, @wlömert2014], we only need to provide the data, the alternatives in the validation/holdout task (\texttt{\color{purple}opts}), and the actual choice (\texttt{\color{purple}choice}). Everything can be implemented in the tidyverse [@tidyverse] logic.

\pagebreak

@@

Finally, we show two functions for market simulations, namely \texttt{\color{purple}marksim} as well as \texttt{\color{purple}turf}. First, we calculate the market shares based on the multinomial logit model [@McFadden1974]. Besides the aggregated shares, \texttt{\color{purple}marksim} also provides standard errors and the 95th confidence interval.

```{r}
marksim(
  data = HOT,
  opts = c(Option_1:None),
  method = "shareofpref"
)
```

\texttt{\color{purple}turf}, a "product line extension model" [@miaoulis1990, p. 29] is a tool to find the perfect assortment that creates the highest reach and a powerful tool for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we could include as well the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, as well as \texttt{\color{purple}prohib}, to define prohibitions of combinations of items that should not be part of the assortment (please see the vignette for more details and also to see how to apply \texttt{\color{purple}turf} with data obtained by a likert scale).

For the following example, let us assume that you conducted an anchored MaxDiff analysis with 10 items (\texttt{\color{purple}opts}) and you want to find the best assortment with a size of 3 (\texttt{\color{purple}size = 3}). As a threshold (\texttt{\color{purple}none}) you use the anchor (no-buy alternative).

>>>>

validateHOT is a package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using [Sawtooth Software](https://sawtoothsoftware.com/).[^1]

[^1]: validateHOT also works with MaxDiff raw logit utilities, see, for example, @chapman2023. For CBC, it should also work with other sofware, as long as all attributes are part-worth coded. Linear and piecewise coding is implemented to work with Sawtooth Software.

Preference measurement techniques', such as (A)CBC or MaxDiff, ultimate goal is to predict future behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and can also predict outside tasks (i.e., the model has external validity). In terms of external validity, we refer to the generalizations to different settings [see, @calder1982, p.240]. The easiest way to test this is to include so-called validation or holdout tasks [@Orme2015], which are tasks that are fixed (i.e., same across participant) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Despite the important role of validation tasks, practitioners often do not include them [@yang2018]. Given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions, the current approach is unsatisfactory.

validateHOT combines both validation and market simulation in one package and has three key advantages: a) it helps you to decide which is the best model to proceed by validating it, b) it runs relevant market simulations that help to find the right product combinations, and finally, c) it is an open source tool including functions that are usually implemented in paid software, and therefore, remain a black-box for researchers and practitioners.

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and therefore ensuring that the model has predictive validity; b) running market simulations (e.g., **T**otal **U**nduplicated **R**each and **F**requency, hereafter TURF). Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix (see \autoref{tab:table1}). However, to put the Sawtooth export into the right format, the user needs some data wrangling which could pose a barrier. Moreover, packages mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR [@ChoiceModelR], choicetools [@choicetools], logitR [@logitr], bayesm [@bayesm] etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is still missing. validateHOT is especially helpful for researchers who plan to report results for research articles according to open science standards. In addition, since the former package turfR is no longer available on the Comprehensive R Archive Network (see [CRAN](https://cran.r-project.org/web/packages/turfR/index.html)), academics and practitioners do not have an alternative function in *R* to run this method. Therefore, practitioners and academics mainly have to stick to paid solutions.

@@

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf} as well as the 2 rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT}, and can be run using the raw logit scores.

| Validation metrics | Confusion matrix | Market simulations | Rescaling scores |
|:----------------:|:----------------:|:----------------:|:----------------:|
|     hitrate()      |    accuracy()    |    freqassort()    |    att_imp()     |
|        kl()        |       f1()       |     marksim()      |  prob_scores()   |
|       mae()        |   precision()    |      reach()       |                  |
|      medae()       |     recall()     |       turf()       |                  |
|       mhp()        |  specificity()   |                    |                  |
|       rmse()       |                  |                    |                  |

@@

After running the Hierarchical Bayes estimation in, for example, Sawtooth Software, the **raw** utility scores have to be exported and afterwards read into *R* as a data frame. This data frame also has to include the actual choice in the validation/holdout task.

To define the validation/holdout task, which has a total of 7 items (\texttt{\color{purple}prod}) plus the no-buy alternative (\texttt{\color{purple}none}), we use the \texttt{\color{purple}createHOT} function. Here, the user can define the attributes as well as the method (here \texttt{\color{purple}MaxDiff}).

@@

Next, to get the relevant validation metrics that are often reported in conjoint studies, for example, hit rate [e.g., @ding2005], mean hit probability [mhp, @voleti2017], or mean absolute error [mae, @wloemert2014], we only need to provide the data, the alternatives in the validation/holdout task (\texttt{\color{purple}opts}), and the actual choice (\texttt{\color{purple}choice}). Everything can be implemented in the tidyverse [@tidyverse] logic.

@@

Finally, we show two functions for market simulations, namely \texttt{\color{purple}marksim} as well as \texttt{\color{purple}turf}. First, we calculate the market shares based on the multinomial logit model [@McFadden1974]. Besides the aggregated shares, \texttt{\color{purple}marksim} also provides standard errors and the 95% confidence interval.

```{r}
marksim(
  data = HOT,
  opts = c(Option_1:None),
  method = "sop"
)
```

\texttt{\color{purple}turf}, a "product line extension model" [@miaoulis1990, p. 29], is a tool to find the perfect assortment that creates the highest reach and especially powerful for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we also include the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, as well as \texttt{\color{purple}prohib}, to define prohibitions of combinations of items that should not be part of the assortment (please see the vignette for more details as well as how to apply \texttt{\color{purple}turf} with data obtained using a likert scale).

For the following example, let us assume that the user conducted an anchored MaxDiff analysis with 10 items (\texttt{\color{purple}opts}) and now wants to find the best assortment with a size of 3 (\texttt{\color{purple}size = 3}). As a threshold (\texttt{\color{purple}none}), the user uses the anchor (no-buy alternative).
